\section{Introduction}

Classification, also known as taxonomy of species, is the scientific process of categorizing and organizing living organisms into a hierarchical system based on their shared characteristics. Taxonomy aims to provide a standardized and universal framework for naming and classifying species. The classification of species involves placing organisms into various ranks or categories based on their similarities and differences. The most fundamental unit of classification is the species, which represents a group of individuals that can interbreed and produce fertile offspring.
The hierarchical system of classification typically follows a structure that starts with the broadest categories and narrows down to more specific groups. The traditional hierarchy consists of eight ranks: Domain, Kingdom, Phylum, Class, Order, Family, Genus, and Species

Traditionally, taxonomy was based on morphology. It relied on observable physical attributes like size, shape, color, and body structure to classify organisms. However, with advancements in molecular biology and DNA sequencing techniques, DNA has become a valuable tool in taxonomy. DNA classification provides a more objective and accurate method for determining relationships between organisms, as DNA carries hereditary information that is passed down from generation to generation.  By using the differences in DNA among different species, scientists can organize and categorize living things based on their genetic makeup in the form of taxonomy or classification system. According to a 2020 study by Sosa et al., morphological analysis and ecological factors should still be considered to improve species recognition and classification (81). Consequently, DNA classification should not be considered the only means of identifying and classifying species. 

Machine learning is an evolving field of technology that has revolutionized the way we approach and solve problems. It involves a system that can learn from data, identify patterns, and make decisions with minimal human intervention. These decisions are more efficient and quicker to process; however, the accuracy of the decisions remains dependent on the data it is exposed to \cite{gama2011}.  Machine learning is increasingly becoming a more popular and reliable method for analyzing and classifying DNA data.  It is used in a variety of ways to classify, analyse and extract useful information from DNA data sets such as automating the segmentation of DNA sequences, which enables greater accuracy and speed of sorting and analysis than traditional methods \cite{chambert2019}. This technology is also used to identify variations in the DNA bases and relate them to medical conditions \cite{Kallman2019}.
Another application is in determining how entire genomes are related to one another across different organisms \cite{cottrell2017}. Also, it is used to identify regions of DNA that are evolutionarily conserved across multiple species \cite{monsalve2019}, or to generate predictions about the function of certain genetic elements. 

Recently several ML algorithms such as Support Vector Machines and Artificial Neural Networks are used to classify features given information about the genetic data, allowing for a more automated approach \cite{kalender2020}. This is especially useful for lab professionals as it reduces the need for manual intervention, allowing for greater efficiency in the classification \cite{vihar2019}. Additionally, due to the speed and accuracy of ML algorithms, scientists can quickly identify changes in genomes and mutations, aiding in research and development of new drugs \cite{neto2021}. In conclusion, ML can be utilised in DNA classification, contributing to the automation and efficacy of the process.

By combining machine learning techniques with other methods, it is possible to classify DNA sequences with higher accuracy and efficiency. This opens up numerous possibilities in the field of genetics, allowing scientists to gain deeper insight into the roles genes and proteins play in disease, development, and many other biological processes.

 


